stupid backoff

### 2019-11-18 al_stup.Rmd
full algorithm

1. create frequency tables and n-grams
```{r}
getFreqs = function(dat, ng) {
        dat.dfm = dfm(dat, ngrams = ng, remove_punct = T, remove_numbers = T,
                      remove = stopwords("english"))
        dat.freq = docfreq(dat.dfm)
        dat.freq = dat.freq[sort(names(dat.freq))] 
        return(dat.freq)
}

@@

```{r}
unigs = getTables(BG, 1)
bigrs = getTables(BG, 2)
trigs = getTables(BG, 3)
```

@@

getODBigs = function(bigPre, bigrams) {
       big.obs = data.frame(ngrams = vector(mode = 'character', length = 0),
                            freq = vector(mode = 'integer', length = 0))
       w_i_1 <- str_split(bigPre, "_")[[1]][2]
       regex = sprintf("%s%s%s", "^", w_i_1, "_")
       indices = grep(regex, bigrams$ngram)
       if(length(indices > 0)) {
               big.obs = bigrams[indices, ]
       }
       return(big.obs)
}

getODUnigs = function(bigPre, unigrams) {

@@

getTrigProbs = function(bigPre, trigrams, bigrams) {

S3 = getODTrigs(bigPre, trigs) # get observed trigrams and counts
S2 = getODBigs(bigPre, bigrs) # get observed bigrams and counts

>>>>

1. create frequency tables and n-grams
```{r}
getFreqs = function(dat, ng) {
        dat.dfm = dfm(dat, ngrams = ng, remove_punct = T, remove_numbers = T
                      #,remove = stopwords("english")
                      )
        dat.freq = docfreq(dat.dfm)
        dat.freq = dat.freq[sort(names(dat.freq))] 
        return(dat.freq)
}

@@

```{r}
BG.unigs = getTables(BG, 1)
BG.bigrs = getTables(BG, 2)
BG.trigs = getTables(BG, 3)
rm(BG)

NS.unigs = getTables(NS, 1)
NS.bigrs = getTables(NS, 2)
NS.trigs = getTables(NS, 3)
rm(NS)

TW.unigs = getTables(TW, 1)
TW.bigrs = getTables(TW, 2)
TW.trigs = getTables(TW, 3)
rm(TW)

@@

## Returns a two column data.frame of observed bigrams that start with the
## unigram prefix (second word of bigPre) in the first column named ngram and
## frequencies/counts in the second column named freq. If no observed bigrams
## that start with second word of bigPre exist, an empty data.frame is returned.
##
## bigPre -  single-element char array of the form w2_w1 which are the first 
##           two words of the trigram we are predicting the tail word of
## trigrams - 2 column data.table. The first column: ngram,
##            contains all the trigrams in the corpus. The second column:
##            freq, contains the frequency/count of each trigram.
getODBigs = function(bigPre, bigrams) {
       big.obs = data.frame(ngrams = vector(mode = 'character', length = 0),
                            freq = vector(mode = 'integer', length = 0))
       w_i_1 <- str_split(bigPre, "_")[[1]][2]
       regex = sprintf("%s%s%s", "^", w_i_1, "_")
       indices = grep(regex, bigrams$ngram)
       if(length(indices > 0)) {
               big.obs = bigrams[indices, ]
       }
       return(big.obs)
}

#x = getTrigProbs(bigPre, trigs, bigrs, unigs)

getTrigProbs = function(bigPre, trigrams, bigrams, unigrams) {
        obs_trigs = getODTrigs(bigPre, trigrams)
        regex = sprintf("%s%s%s", "^", bigPre, "$")
        bigr_count = bigrams[grep(regex, bigrams$ngram), ]$freq

        if(nrow(obs_trigs) < 1) {
                bigr_probs = getBigProbs(bigPre, bigrams, unigrams)
                probs = data.frame(ngram = bigr_probs$ngram, 
                                   prob = bigr_probs$prob * 0.4)
                return(probs)
        } else {
                probs = obs_trigs %>% rename( prob = freq) %>%
                    mutate(prob = prob / bigr_count) %>% 
                    arrange(desc(prob), ngram)
                return(probs)
        }

@@

getBigProbs = function(bigPre, bigrams, unigrams) {
    obs_bigs = getODBigs(bigPre, bigrams)
    w_i_1 <- str_split(bigPre, "_")[[1]][2]
    regex = sprintf("%s%s%s", "^", w_i_1, "$")
    unig_count = unigrams[grep(regex, unigrams$ngram), ]$freq
    if(nrow(obs_bigs) < 1) {
        probs = getUnigProbs(unigrams) %>% mutate(prob = prob * 0.4)
        return(probs)
    } else {
        probs = obs_bigs %>% rename(prob = freq) %>%
            mutate(prob = prob/unig_count) %>%
            arrange(desc(prob), ngram)
        return(probs)
    }
}

getUnigProbs = function(unigrams) {
    probs = unigrams %>% rename(prob = freq) %>%
        mutate(prob = prob/sum(prob)) %>%
        arrange(desc(prob), ngram)
    return(probs)
}
